### 2023-03-17 chapter0/chapter0.tex
Finished SLiM paragraph!

The structural diversity of IDRs is directly related to their functional plasticity, and as IDRs are not confined to one conformation, they can interact with and bind to many possible partners. Often these partners are other proteins, but they can also be other macromolecules like DNA or even small molecules and ions. As a result of this adaptability, IDRs are enriched in proteins involved in cell signaling and regulation. Because cells are highly compartmentalized, they have a dizzying array of mechanisms to relay messages.\footnote{The most fundamental compartment is the cell, which roughly separates inside from outside and life from non-life. However, the cells of more complex organisms called eukaryotes contain additional subcompartments called organelles. Compartmentalization is essential in living systems because it confines and concentrates biochemical reactions that would be harmful if they occurred at the wrong place or time. However, it also introduces many complications because information in the form of physical or chemical signals cannot travel freely.} Many are mediated by the interaction of one protein with another, which in turn creates a change in the state of the system that propagates the signal. Often these state changes are chemical alterations made by one protein to another, which are termed ``post-translational modifications'' (PTMs).\footnote{A protein's amino acid sequence is encoded in a cell's DNA, and the process of reading that information to create a protein (from an intermediate molecule called RNA) is translation. Therefore, any modifications to a protein after its initial synthesis are post-translational.} One of the most common modifications is phosphorylation where a phosphate group is attached to a specific amino acid in the protein. Phosphate groups contain three negative charges, so their addition can dramatically affect a protein's structural energetics and induce a conformational change. Thus, phosphorylation often plays a key role in toggling proteins between inactive and active states. In general, though, PTMs encompass a variety of chemical modifications with similarly diverse impacts on a protein's behavior. Because IDRs are generally exposed to their environment, they are frequent targets of PTMs. However, PTMs do not occur haphazardly in proteins but are instead targeted to binding sites created by sequential patterns of residues called short linear motifs (SLiMs). SLiMs are a general mechanism for mediating interactions with proteins, so while many SLiMs are targets of PTMs, others simply recruit binding partners. In contrast to the highly structured interfaces that characterize folded domains, however, SLiMs are usually no more than ten residues. Thus, their interactions are relatively weak and therefore highly transient. As their flexibility makes SLiMs easily accessible to binding partners, they are highly enriched in IDRs.

SLiMs can also occur combinatorially.
Furthermore, SLiMs are easily tuned because PTMs or evolution can change their properties by modifying only a few residues. This dynamism and modularity is essential in signalling as it allows the efficient propagation and modulation of information.
Thus, they often function as hubs in networks of signaling pathways [CITATION]
Some IDRs are roughly like molecular computers which integrate signals from varying sources
As IDRs are highly flexible, they can bind to many partners, sometimes simultaneously, and therefore often act as hubs in these signaling networks.

>>>>

The structural diversity of IDRs is directly related to their functional plasticity, and as IDRs are not confined to one conformation, they can interact with and bind to many possible partners. Often these partners are other proteins, but they can also be other macromolecules like DNA or even small molecules and ions. As a result of this adaptability, IDRs are enriched in proteins involved in cell signaling and regulation. Because cells are highly compartmentalized, they have a dizzying array of mechanisms to relay messages.\footnote{The most fundamental compartment is the cell, which roughly separates inside from outside and life from non-life. However, the cells of more complex organisms called eukaryotes contain additional subcompartments called organelles. Compartmentalization is essential in living systems because it confines and concentrates biochemical reactions that would be harmful if they occurred at the wrong place or time. However, it also introduces many complications because information in the form of physical or chemical signals cannot travel freely.} Many are mediated by the interaction of one protein with another, which in turn creates a change in the state of the system that propagates the signal. Often these state changes are chemical alterations made by one protein to another, which are termed ``post-translational modifications'' (PTMs).\footnote{A protein's amino acid sequence is encoded in a cell's DNA, and the process of reading that information to create a protein (from an intermediate molecule called RNA) is translation. Therefore, any modifications to a protein after its initial synthesis are post-translational.} One of the most common modifications is phosphorylation where a phosphate group is attached to a specific amino acid in the protein. Phosphate groups contain three negative charges, so their addition can dramatically affect a protein's structural energetics and induce a conformational change. Thus, phosphorylation often plays a key role in toggling proteins between inactive and active states. In general, though, PTMs encompass a variety of chemical modifications with similarly diverse impacts on a protein's behavior. Because IDRs are generally exposed to their environment, they are frequent targets of PTMs. However, PTMs do not occur haphazardly in proteins but are instead targeted to binding sites created by sequential patterns of residues called short linear motifs (SLiMs). SLiMs are a general mechanism for mediating interactions with proteins, so while many SLiMs are targets of PTMs, others simply recruit binding partners. In contrast to the highly structured interfaces that characterize folded domains, however, SLiMs are usually no more than ten residues. Thus, their interactions are relatively weak and highly transient. As their flexibility makes SLiMs easily accessible, they are also enriched in IDRs, which in turn allows them to bind to many partners, sometimes simultaneously. IDRs therefore often act as hubs in complex regulatory networks by propagating signals from diverse sources or organizing binding partners into higher-order structures. Furthermore, these signals and interactions are easily tuned because PTMs can modulate their meanings and strengths, respectively, by modifying only a few residues. Thus, IDRs are like ``molecular computers'' that integrate complex data and respond accordingly to execute different genetic ``programs.''
